Hysteresis in human computation:\\ how one task affects another

### 2015-01-20 docs/publications/2014.09_SCIENCE/man.tex
work on manuscript, add hypothesis tests, put dictionaries in repo

\usepackage{framed, color}
\usepackage{soul}
\definecolor{blu}{rgb}{0,0,1}
\newcommand{\td}[1]{{\color{blu}\hl{TODO: #1}}}
\usepackage{graphicx}
\usepackage{amsthm}
\newtheorem{mydef}{Definition}
\usepackage{dcolumn}
\usepackage{multirow}
\usepackage{booktabs}
\newcolumntype{d}{D{.}{.}{4.0}}
\newcolumntype{s}{D{.}{.}{1.4}}

@@

While microtask platforms have been widely adopted as a research tool,
computer scientists also hold such platforms as the object of 
research, viewing  microtask platforms as a new kind of computing 
architecture.  In analogy to CPUs (central 
processing units) and GPUs (graphics processing units), 
Davis \textit{et al} invoked the \textit{human processing unit}, or HPU \cite{5543192}.
Researchers have sought to define a basic HPU instruction set, and
libraries for HPU programming are under active development
 \cite{little2010turkit,minder2011crowdlang,minder2012crowdlang,kittur2011crowdforge}.

Given the increasing prevalence of microtask-based methods, there
is a desire to understand factors affecting the reliability of microtask
platforms.  Researchers have investigated task-design parameters such as the 
level of pay \cite{kazai2013analysis}, the training and pre-screening of 
workers \cite{le2010ensuring,paolacci2010running}, and user-interface 
design \cite{Finnerty2013}.  
Researchers have also investigated the effects of \textit{framing}, 
by altering the description of the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, and the problem description
\cite{thibodeau2013natural}.

@@

Researchers investigating a somewhat similar class of psychological phenomena,
found that ratings that microtask workers give to items in a list
depend on the order in which the list is presented, an effect known as
\textit{positional bias} \cite{lerman2014leveraging}.  While this 
result applies mainly to tasks of a 
particular design, it shows that basic psychological phenomena
can profoundly influence crowdsourced data.  If human computation is to be
understood as a kind of computing architecture, it will certainly be 
necessary to account for the computational ramifications of these 
psychological phenomena at play in the HPU.

The results we present here show that a worker's response during a task
does indeed depend strongly on the tasks performed beforehand.   
In engineering terms, 
unlike CPUs and GPUs, HPUs are subject to \textit{hysteresis}, meaning
that their outputs depend on the history of their inputs.  Because 
task-repetition is inherent in their design, virtually all microtask studies,
and most crowdsourced studies are at risk from the form of bias we 
investigate here. We call the effects that earlier tasks exert on later ones 
\textit{intertask effects}.

We investigated intertask effects using image-labeling tasks on the Amazon 
Mechanical Turk (MTurk) microtask platform.  Image-labeling tasks are among
the most common MTurk tasks 
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running}. 
This trend is likely to continue 
in the long run, because image labeling is still very difficult to automate
\cite{5543192}, and is used in many applications, such as website moderation, 
computer vision research, and social media analytics (\# insert other refs) 
\cite{5543192}.  In our first experiment, which we will refer to as 
\textit{intertask-food-objects}, workers were assigned to either a
\textit{food} or \textit{objects} treatment.  Workers in these treatments
performed a set of five \textit{initial tasks}, which each involved 
providing five descriptive labels for an image depicting 
either food or (non-food) objects, depending on the worker's 
treatment (see Fig.~\ref{fig:task}A and B).  Following the initial 
tasks, workers performed five \textit{test tasks}, which were identical for 
both treatments.  The test tasks contained images depicting a 
mixture of food and objects (see Fig.~\ref{fig:task}C), and workers
were again asked to provide five labels.  Five initial 
tasks, together with the test tasks, comprised one HIT 
(\# workers were only allowed to participate once). During the HIT, workers 
were presented images one at a time, and, crucially, there was no distinction
or interruption between the initial and test tasks.

@@

While this affirms the \textit{existence} of intertask effects, we wish to
understand the nature and severity of the bias induced by prior task 
exposures.  To this end, we define the extent of bias, $\theta$, induced by 
alternative exposures, to 
be the resulting change in the probability with which workers provide given 
responses (\# this is a general divergence metric between probability 
distributions known as the \textit{total variation distance}):
\begin{equation}
	\theta = \frac{1}{2}\sum_{x \in X} \left| p_1(x) - p_2(x) \right|,
\end{equation}
where $p_i(x)$ is the probability that a worker subjected to treatment $i$
provides response $x$, from the set of possible responses $X$.
It is important to emphasize that a significant result from a $\chi^2$ test 
only allows us to conclude that $\theta$ is non-zero.  Determining $\theta$ from finite samples of 
responses, using purely statistical approaches, is difficult (\#), which 
leads us to reformulate the
problem in terms of classification: given a worker's 
response, can a machine reliably infer what her prior task-exposure had 
been?  Intuitively, the greater the effect that prior tasks have on subsequent
responses, the more accurate the classifier can be.  Formally, as we show
in the supplementary text, the accuracy of such a classifier, $\eta$,
yields a lower bound on the bias:
\begin{equation}
	\theta \geq 2\eta - 1.
	\label{l1}
\end{equation}
By measuring the accuracy of a classifier, we can bound the 
bias induced by any perturbation, including both intertask effects and 
framing.  As defined, bias ranges from 0 (meaning the exposures 
do not affect the probabilities of responses), to 1 
(or 100\%) (meaning that under different exposures, workers provide completely
different sets of responses).

@@

We conducted variants of these experiments using different images, to see 
whether this trend was robust.  In the experiment 
\textit{intertask-food-culture},
workers were either assigned to a \textit{food} or \textit{culture} treatment.
As before, the initial tasks for the \textit{food} treatment contained images 
depicting food (see Fig.~\ref{fig:task}D). 
In the \textit{culture} treatment, 
the initial tasks contained images depicting cultural scenes 
(of dance, sport, or music) (see Fig.~\ref{fig:task}E).  The test tasks, which were identical for both
treatments, included images depicting meals of identifiable cultural 
origin 
(see Fig.~\ref{fig:task}F).  Results for this experiment again showed a 
strong bias as a result of intertask effects (about 50\%) 
(see Fig.~\ref{fig:theta}A).  Again, we compared intertask effects to 
framing, by performing an experiment (\textit{frame-food-culture}) using 
the same test tasks as in \textit{intertask-food-culture}, in which we exposed
workers to a statement framing the purpose of the tasks as the recognition 
of either food or culture.  
For this framing experiment, we did not observe a statistically significant 
difference between the labels attributed by the \textit{food} and 
\textit{culture} treatments during test tasks (see Fig.~\ref{fig:theta}).

@@

Prior to this study, little thought appears to have been given to the 
ordering and bundling of microtasks.  But our findings show that these are 
powerful design factors.  A natural response might be simply to randomize 
task ordering, a practice that is commonly employed.  But the sheer extent of 
bias that we observed leads us to question 
whether that is the best approach.  Although the bias would be random in 
that case, our results suggest that it is on the order of 30-50\%, 
which is a major contribution to noise in the results.
It would be preferable to understand intertask
effects in greater depth so that they might \textit{controlled} rather than
simply randomized.

@@

@@

The results show that good inter-rater reliability was achieved during the
validation (82.4\%).  Looking at the agreement between the 
machine codes and the majority human-generated codes, we find they agree
88\% of the time.  This means that the concept used to instruct human coders
provides a good approximation to what the wordnet-based machine codes actually
represent. In other words, the machine codes correspond closely to indicating
which words correspond to nouns that refer to edible things, and which ones
don't, which provides a clear and simple interpretation for the machine
coding of food and non-food words.

@@

\subsection*{S9: Justifications and statistical properties of classifier-based measurement of computational hysteresis}
\begin{itemize}
	\item{Relationship to L1-Distance}
	\item{Relationship to bias}
	\item{Relationship to Jenson-Shannon divergence}
	\item{Other approaches to measuring L1-distance}
	\item{Comparison to hypothesis testing using $\chi^2$}
\end{itemize}

>>>>

\usepackage{amsmath}
\usepackage{framed, color}
\usepackage{soul}
\definecolor{blu}{rgb}{0,0,1}
\newcommand{\td}[1]{{\color{blu}\hl{TODO: #1}}}
\usepackage{graphicx}
\usepackage{amsthm}
\newtheorem{mydef}{Definition}
\usepackage{dcolumn}
\usepackage{multirow}
\usepackage{booktabs}
\newcolumntype{d}{D{.}{.}{4.0}}
\newcolumntype{s}{D{.}{.}{1.4}}

@@

These platforms allow requesters to make use of human intelligence
\textit{algorithmically}.  Tasks are posted and results fetched by transfering
files over the Internet, making it possible to embed microtask work in an
automated process.  To the surrounding process, workers seem like 
input-output devices, which provides much of the flexibility and 
cost-savings of actually automating the work: the workforce is available
on demand, over the Internet, without any contracts or interviews, at a rate
of about 6 USD\$/ hr.  Many kinds of work can be performed at a fraction of 
the cost that would otherwise be required, either to recruit temporary 
workers, or to research and design a sophisticated automated solution.

For this reason, there has been a surge in demand for microtask work
in both industry and academia.  Industry employs microtask workers
for clerical tasks, such as digitizing records and correcting errors; 
and for marketing activities, such as running focus groups, writing product 
reviews,
and moderating website content.  Academia turns to microtask workers as 
a way to recruit study participants, and to annotate datasets, for example,
to provide ground-truth in training artificial intelligence systems.

Given the increasing prevalence of microtask-based methods, there
is a desire to understand the factors affecting the reliability of microtask
platforms.  Researchers have investigated task-design parameters such as the 
level of pay \cite{kazai2013analysis}, the training and pre-screening of 
workers \cite{le2010ensuring,paolacci2010running}, and user-interface 
design \cite{Finnerty2013}.  
Researchers have also investigated the effects of \textit{framing}, 
by altering the description of the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, and the problem description
\cite{thibodeau2013natural}.

@@

Researchers investigating a somewhat similar class of psychological phenomena,
found that ratings that microtask workers give to items in a list
depend on the order in which the list is presented, an effect known as
\textit{positional bias} \cite{lerman2014leveraging}.  This serves as an 
example of a basic psychological phenomenon impacting crowdsourced data,
and raises a serious question about what other cognitive biases might
affect microtask work, and how prevelant induced bias might be.

We therefore turn to the question of bias induced by taks-repetition effects, 
which, if present, would impact virtually every microtask workflow.
The results we present here show that a worker's response during a task
does indeed depend strongly on the tasks performed beforehand.   
We call the effects that earlier tasks exert on later ones 
\textit{intertask effects}.

As we have described, microtask platforms have been widely adopted as a 
tool, but computer scientists also hold such platforms as the object of 
research, viewing  microtask platforms as a new kind of computing 
architecture.  In analogy to CPUs (central 
processing units) and GPUs (graphics processing units), 
Davis \textit{et al} invoked the \textit{human processing unit}, or HPU \cite{5543192}.
Researchers sought to define a basic HPU instruction set, and
libraries for HPU programming are under active development
 \cite{little2010turkit,minder2011crowdlang,minder2012crowdlang,kittur2011crowdforge}.  
HPU outputs are inherently stochastic, but our revelation of strong intertask 
effects shows that HPUs are subject to \textit{hysteresis}, 
meaning that their 
outputs depend on the history of their inputs.  Reproducibility and 
path-independance are crucial for properties for the basic components in a 
computing architecture, so it will be necessary to understand the 
computational ramifications of these psychological phenomena to further this 
interesting line of investigation.

We investigated intertask effects using image-labeling tasks on the Amazon 
Mechanical Turk (MTurk) microtask platform.  
In addition to being one of the most common kinds of tasks
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running},
image-labeling is prototypical for any kind of task in which the worker is 
asked to provide a qualitative free-text response to a prompt.  
This includes such such examples as describing
video and audio recordings, summarizing texts, or providing personal 
experiential feedback to a product review.
%in the long run, because image labeling is still very difficult to automate
%\cite{5543192}, and is used in many applications, such as website moderation, 
%computer vision research, and social media analytics (\# insert other refs) 
%\cite{5543192}.

In our first experiment, which we will refer to as 
\textit{intertask-food-objects}, workers were assigned to either a
\textit{food} or \textit{objects} treatment.  Workers in these treatments
performed a set of five \textit{initial tasks}, which each involved 
providing five descriptive labels for an image depicting 
either food or (non-food) objects, depending on the worker's 
treatment (see Fig.~\ref{fig:task}A and B).  Following the initial 
tasks, workers performed five \textit{test tasks}, which were identical for 
both treatments.  The test tasks contained images depicting a 
mixture of food and objects (see Fig.~\ref{fig:task}C), and workers
were again asked to provide five labels.  Five initial 
tasks, together with the test tasks, comprised one HIT 
(\# workers were only allowed to participate once). During the HIT, workers 
were presented images one at a time, and, crucially, we made no 
distinction between the initial and test tasks.

@@

While this affirms the \textit{existence} of intertask effects, we wish to
understand the nature and severity of the bias induced by prior task 
exposures.  To this end, we define the extent of bias, $\theta$, in terms of
the strength of influence that a perturbation has on worker's responses.
Since different workers generally provide different responses for the same 
task, we characterize worker's responses by a probability distribution.  
The extent of bias is then defined by the difference in the response 
distribution for different treatments of workers
(\# this is a general divergence metric between probability 
distributions known as the \textit{total variation distance}):
\begin{equation}
	\theta = \frac{1}{2}\sum_{x \in X} \left| p_0(x) - p_1(x) \right|,
	\label{eq:theta}
\end{equation}
where $p_i(x)$ is the probability that a worker subjected to treatment $i$
provides response $x$, from the set of possible responses $X$.
It is important to emphasize that a significant result from a $\chi^2$ test 
only allows us to conclude that $\theta$ is non-zero.  Determining $\theta$ from finite samples of 
responses, using purely statistical approaches, is difficult (\#), which 
leads us to reformulate the
problem in terms of classification: given a worker's 
response, can a machine reliably infer what her prior task-exposure had 
been?  Intuitively, the greater the effect that prior tasks have on subsequent
responses, the more accurate the classifier can be.  Formally, as we show
in the supplementary text, the accuracy of such a classifier, $\eta$,
yields a lower bound on the bias:
\begin{equation}
	\theta \geq 2\eta - 1.
	\label{l1}
\end{equation}
By measuring the accuracy of a classifier, we can bound the 
bias induced by any perturbation, including both intertask effects and 
framing.  As defined, bias ranges from 0 (meaning the exposures 
do not affect the probabilities of responses), to 1 
(or 100\%) (meaning that under different exposures, workers provide completely
different sets of responses).

@@

We conducted variants of these experiments using different images, to see 
whether this trend was robust.  In the experiment 
\textit{intertask-food-culture},
workers were either assigned to a \textit{food} or \textit{culture} treatment.
As before, the initial tasks for the \textit{food} treatment contained images 
depicting food (see Fig.~\ref{fig:task}D). 
In the \textit{culture} treatment, 
the initial tasks contained images depicting cultural scenes 
(of dance, sport, or music) (see Fig.~\ref{fig:task}E).  The test tasks, which were identical for both
treatments, included images depicting meals of identifiable cultural 
origin (\# the initial tasks in 
\textit{intertask-food-culture} are arguably also of identifiable cultural 
origin, but this would tend to reduce, rather than inflate, our results
(we discuss this further in the supplementary material).)
(see Fig.~\ref{fig:task}F).

Results for this experiment again showed a 
strong bias as a result of intertask effects (about 50\%) 
(see Fig.~\ref{fig:theta}A).  Again, we compared intertask effects to 
framing, by performing an experiment (\textit{frame-food-culture}) using 
the same test tasks as in \textit{intertask-food-culture}, in which we exposed
workers to a statement framing the purpose of the tasks as the recognition 
of either food or culture.  
For this framing experiment, we did not observe a statistically significant 
difference between the labels attributed by the \textit{food} and 
\textit{culture} treatments during test tasks (see Fig.~\ref{fig:theta}).

@@

Prior to this study, little thought appears to have been given to the 
ordering and bundling of microtasks.  But our findings show that these are 
powerful design factors.  A natural response might be simply to randomize 
task ordering, a practice that is commonly employed.  But the sheer extent of 
bias suggests that this will still introduce a significant amount of noise
into the results.  Even chains of two or three similar tasks, which
will not be reliably eliminated in random permutations, could lead
to the levels of bias we observed in our experiments.
It would be preferable to understand intertask
effects in greater depth so that they might be properly \textit{controlled}.

@@

\subsection*{Determination of $\theta$}

@@

The results show that good inter-rater reliability was achieved during the
validation (82.4\%).  Looking at the agreement between the 
machine codes and the majority human-generated codes, we find they agree
88\% of the time.  This means that the concept used to instruct human coders
provides a good approximation to what the wordnet-based machine codes actually
represent. In other words, the machine codes correspond closely to indicating
which words correspond to nouns that refer to edible things, and which ones
do not, which provides a clear and simple interpretation for the machine
coding of food and non-food words.

@@

\paragraph{Statistical significance and confidence intervals for $\theta$.}
The measurements of $\theta$ were based on the number of correct 
classifications during validation tests.  We used leave-one-out validation,
and count the number of correct classifications, $X$, which has a binomial
distribution.  We take the null hypothesis that the classifier performs
no better than chance, $X \sim \mathrm{Bin}(N,0.5)$, where $N$ is the number 
of workers in both treatments of an experiment.  Then, the critical number 
of correct classifications, $x^*$, in for a one-tailed hypothesis test with
significance $\alpha$ is:
\begin{align}
	&x^* = \sup \left\{
			x: \mathrm{Pr}\{ \mathrm{Bin}(N;0.5) \geq x \} < \alpha
		\right\}\\
	\text{where}\quad &\mathrm{Pr}\{ \mathrm{Bin}(N,p) \geq x \} = 
		\sum_{x'=x}^{N} \binom{N}{x'}p^{x'}(1-p)^{(N-x')}
\end{align}
For the treatments in \textit{intertask-food-objects}, 5 different 
permutations of each treatment was tested, and a different classifier was 
built for each permutation.  For the sake of estimating the accuracy of 
these classifiers, we assume that they have the same accuracy, which means
that \textit{intertask-food-objects} has a larger number of validation tests
(larger $N$), and for this hypothesis test we assume that $X$ has a normal 
distribution, with $\mu=N/2$ and $\sigma = \frac{1}{2}\sqrt{N}$.

To determining the confidence intervals for measured $\theta$ values we 
use the exact Clopper-Pearson method.  
Based on the observed number of successes, $X$, we calculate 
$\eta^*_\mathrm{low}$ and $\eta^*_\mathrm{high}$ from:
\begin{align}
	\eta^*_\mathrm{high} &= \sup
		\left\{
			\eta : \mathrm{Pr}\{\mathrm{Bin}(N; \eta) \leq X \} > 
				\frac{\alpha}{2}
		\right\} \\
	\eta^*_\mathrm{low} &= \inf
		\left\{
			\eta : \mathrm{Pr}\{\mathrm{Bin}(N; \eta) \geq X \} > 
				\frac{\alpha}{2}
		\right\}
\end{align}
Again, for the experiment \textit{intertask-food-objects}, which had 
more validation tests than the others, we derive the confidence intervals
assuming $X$ is normally distributed.
Since the significance values are to be shown on a plot of $\theta$, we 
transform them into equivalent values of 
$\theta$: $\theta^*_\mathrm{high} = 2\eta^*_\mathrm{high} - 1$ and 
$\theta^*_\mathrm{low} = 2\eta^*_\mathrm{low} - 1$.
The statistics collected for the measurement of $\theta$ are tabulated in 
table X.

\begin{table}
\begin{center}
\begin{tabular}{c c c c c c c}
	\toprule
	Experiment & $N$ & $X$ & $x^*$ &
		$\hat{\theta}$ & $\theta^*_\mathrm{low}$ & 
		$\theta^*_\mathrm{high}$ \\
	\midrule
	\textit{intertask-food-objects} & 1190 & 777 & 624 & \textbf{30.6} 
		& 25.2 & 36.0 \\
	\textit{frame-food-objects} & 238 & 122 & 133 & 2.5 & -10.6 & 14.7 \\
	\textit{echo-food-objects} & 238 & 162 &  133 & \textbf{36.1} & 23.4 
		& 47.1  \\
	\textit{intertask-food-culture} & 238 & 180 & 133 & \textbf{51.3} & 39.3 
		& 61.1 \\
	\textit{frame-food-culture} & 238 & 130 & 133 & 9.2 & -3.9 & 21.3 \\
	\bottomrule
\end{tabular}
\caption{Statistics for the measurement of the extent of bias, $\theta$,
	induced by intertask and framing effects in various experiments.
	Number of validation tests, $N$; number of successful classifications, 
	$X$; critical number of succesful classifications to reject the null 
	hypotheses that the classifier does no better than chance 
	(one-tailed test), $x^*$; 
	lower bound estimate of bias based on observed number of successful
	classifications, $\hat{\theta}$; lower confidence interval for lower
	bound estimate of bias, $\theta^*_\mathrm{low}$; upper confidence 
	interval for same, $\theta^*_\mathrm{high}$.  Hypothesis test and 
	confidence intervals are based on a significance of $\alpha=0.05$.
	Values for $\hat{\theta}$ in boldface are significantly greater than 0.
}

\end{center}
\end{table}

\subsection*{S9: Justifications and statistical properties of classifier-based measurement of computational hysteresis}

\paragraph{Comparing $\theta$ to a $\chi^2$ significance test}
Suppose that we have a very slightly unfairly weighted coin, which turns
up heads ($x=\mathbf{H}$) more often than expected by 1 part in one hundred:
$\Pr\{x=\mathbf{H}\} = 0.500001$. Such a coin would be fine for many 
practical purposes, like deciding who buys lunch.  We could
perform a $\chi^2$ test to determine if the coin is biased, by flipping 
it 100 million times, while doing the same with a coin we know is fair.
Simulating this with a random number generator yields the following results:

\begin{center}
\begin{tabular}{c c c c}
\toprule
coin & $N_\mathbf{H}$ & $N_{\mathbf{T}}$ & total \\
\toprule
\textit{fair} & 50002283 & 49997717 & $10^8$\\
\textit{biased} & 50101115 & 49898885 & $10^8$ \\
\bottomrule
total & 100103398 & 99896602 & $2\times 10^8$ \\
\bottomrule
\end{tabular}
\end{center}
A $\chi^2$ test yields the following results:
\begin{center}
	\begin{tabular}{c c c c }
	\toprule
	Degrees of freedom & $\chi^2$ & $p$-value \\
	\toprule
	1 & 195.4 & $2.2 \times 10^{-44}$ \\
	\bottomrule
	\end{tabular}
\end{center}
Even though the $\chi^2$ test shows, with tremendous statistical 
significance, that the unfair coin is indeed unfair, the actual extent
of bias is small, 
$\theta = \frac{1}{2}\left( |0.501 - 0.499| + |0.499 - 0.501| \right) = 0.002$.
A sufficiently large sample size will lead to the detection an arbitrarily 
small bias with arbitrarily high statistical significance.
This is why to measure the practical significance of intertask effects
we rely on a measure of effect size, which is in this case a measure
of statistical divergence, $\theta$.

\paragraph{Statistical approaches to measuring bias.}
Recall that in Eq.~\ref{eq:theta} in the main text, we defined the
extent of bias that exists between two groups of workers to be $\theta$:
$$
	\theta = \frac{1}{2}\sum_{x \in X} \left| p_0(x) - p_1(x) \right|,
$$
where $p_0(x)$ and $p_1(x)$ represent the probability mass functions
for responses from two differently treated populations of workers.  
Let us denote these two populations by $P_0$ and $P_1$.
As mentioned, this is a measure of the divergence between $p_0$ and $p_1$ 
which is called either the \textit{total variation} or the 
\textit{L1-distance}.

Measuring the \textit{L1-distance} between two probability distributions,
given only samples, is difficult.  The na\"ive approach is to take
the maximum likelihood estimate of $p(x)$, which sets it equal to the 
frequency with which the response $x$ has been observed, 
$\hat{p}(x) = \frac{N_x}{N}$, and then estimate $\theta$ 
using these frequencies: 
$\hat{\theta} = \sum_{x \in X}|\hat{p}_1(x) - \hat{p}_2(x)|$.  
However, this method tends to drastically overestimate $\theta$.
Recently progress has been made on this problem, and two algorithms have 
been proposed to estimate the L1-distance between distributions
given only samples.  But these methods provide only asymptotic guarantees
which cannot be directly applied to our case.  However, machine learning
approaches provide a general alternative that we describe in the next 
section.

\paragraph{Using classifiers to measure bias.}

We will now show that, any classifier algorithm, $\mathcal{A}$, that
takes the response, $x$, of a worker, and guesses the population to which
the worker belongs ($P_0$ or $P_1$), will do so with an accuracy, $\eta$,
that is bounded according to Eq.~\ref{l1}, which is reproduced here:
$$
	\theta \geq 2\eta - 1.
$$

To do so, we will first formalize $\eta$ by defining a 
\textit{validation test}.  In a validation test, a uniform random bit 
$z\in\{0,1\}$ is sampled, and then, according to the value of $z$, the
response of a worker, $x$, is sampled uniformly randomly from $P_z$.  
The algorithm is provided
with $x$ and yields a guess, $b=\mathcal{A}(x)$, as to whether the worker was 
from $P_0$ or $P_1$.  We shall denote such a validation test by 
$V(P_0, P_1, \mathcal{A})$, and define the value of the test to be 1 if 
$b=z$ and zero otherwise.

We then define the accuracy of classifier $\mathcal{A}$ to be $\eta$ according
to:
$$
\eta = \mathrm{E\{V(P_0, P_1, \mathcal{A})\}} 
	= \mathrm{Pr}\{V(P_0, P_1, \mathcal{A})=1\}.
$$

To develop the relationship between $\eta$ and $\theta$, let us consider
the accuracy $\eta^*$ of the best possible classifier $\mathcal{A}^*$.
The best possible classifier will not in general have perfect accuracy:
if the responses from workers in $P_0$ and $P_1$ are distributed 
identically, then there is no hope of determining which population the worker 
came from given only her response $x$.  But, assuming that responses are
not distributed identically, then given some $x$, the best possible 
classifier must always guess 1 whenever $p_1(x) > p_0(x)$ and otherwise 
guess 0.

Sacrificing, for a moment, some generality, let us assume that, for a given 
$x$, $p_1(x) > p_0(x)$.  
Then, given some $x'$ has been chosen during a 
validation test, the optimal classifier must guess that 
$z = \arg\!\max_{z'}(p_{z'}(x)) = 1$.
The probability that the classifier guesses correctly, given a validation
test in which the resonse $x$ was chosen, is therefore:
\begin{align}
	\mathrm{Pr}\{V = 1 | x' = x \} 
		&= \mathrm{Pr}\{\mathcal{A^*}(x) = z | x' = x \} \\
		&= \mathrm{Pr}\{z = 1 | x' = x \}  \\
		&= \frac{\mathrm{Pr}\{z = 1 , x' = x\}}
			{ \mathrm{Pr}\{z=0 , x'=x\} + \mathrm{Pr}\{z=1 , x'=x\}} \\
		&= \frac{p_1(x)}{p_0(x) + p_1(x)}.
\end{align}
And now, with full generally,
\begin{align}
	\mathrm{Pr}\{V=1 | x' = x \} 
		&= \mathrm{Pr}\{\mathcal{A^*}(x) = z | x' = x \} \\
		&= \mathrm{Pr}\{z = \arg\!\max_{z'}\big(p_{z'}(x)\big)| x' = x \}  \\
		&= \frac{\max\big( p_0(x),p_1(x) \big)}
		{ p_0(x) + p_1(x) }.
\end{align}
We can rewrite:
\begin{align}
	\max\big(p_0(x),p_1(x)\big) = \frac{1}{2}
		\big(
			p_0(x) + p_1(x) + |p_0(x) - p_1(x)|
		\big),
\end{align}
so,
\begin{align}
	\mathrm{Pr}\{V = 1 | x' = x \} 
	&= \frac{1}{2} + \frac{|p_0(x) - p_1(x)|}{2 \big(p_0(x) + p_1(x) \big)}.
\end{align}
Then the overall accuracy that probability that $\mathcal{A^*}$ guesses correctly in 
any validation test is:
\begin{align}
	\eta^* &= \mathrm{Pr}\{V=1\} \\
		&=\sum_{x\in X} \mathrm{Pr}\{V = 1 | x' = x \} \mathrm{Pr}\{x' = x\}\\
		&= \sum_{x\in X} \left(
				\frac{1}{2} + 
				\frac{|p_0(x) - p_1(x)|}{2 \big(p_0(x) + p_1(x) \big)}
			\right)
			\left(
				\frac{p_0(x) + p_1(x)}{2}
			\right)\\
		&= \frac{1}{2}\left(
				\frac{1}{2}(
					\sum_{x \in X}p_0(x) + \sum_{x \in X}p_1(x)
				) + 
				\frac{1}{2} \sum_{x \in X} \big(|p_0(x) + p_1(x)|\big)
			\right) \\
		&= \frac{1}{2} \left( \frac{1}{2}(1 + 1) + \theta \right)\\
		&= \frac{1 + \theta}{2}\\
		\implies \theta &= 2\eta^* - 1.
\end{align}
By definition, no classifier can be more accurate than $\mathcal{A^*}$, so
for every $\mathcal{A}$
\begin{align}
		\theta \geq 2\eta - 1.
\end{align}
